|

### 2022-07-20 manuscript_BE/manuscript_BE_word.Rmd
New version of the manuscript

---
title: Field and experimental data together with computational 
  models reveal how cleaner fish adjust decisions  in a biological market
author:
  - name: Andrés E. Quiñones 
    email: andresqp@gmail.com
    affiliation: a
    corresponding: andresqp@gmail.com
  - name: Zegni Triki
    email: zegni.triki@gmail.com
    affiliation: b
  - name: Redouan Bshary
    email: redouan.bshary@unine.ch
    affiliation: Institute of Biology, University of Neuchatel
address:
  - code: a
    address: Institute of Biology, University of Neuchatel, Rue Emile-Argand 11, 2000-CH Neuchatel, Switzerland
  - code: b
    address: Department of Zoology, Stockholm University, Svante Arrheniusväg 18 B,  106 91 Stockholm, Sweden,
abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. We illustrate this point by fitting computational models to experimental data on performance by wild-caught cleaner fish *Labroides dimidiatus* in an ephemeral reward task, as well as cleaner and client fish densities from the locations of capture. Using Bayesian statistics to fit the model parameters to performance data revealed that cleaner fish most likely estimate future consequences of an action, while it appears unlikely that the removal of the ephemeral reward acts as psychological punishment (negative reinforcement). Incorporating future consequences also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement mechanism. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinnings of cognitive performance.

author_summary: |
  Performance in behavioural experiments is often used to assess the cognitive abilities of animals. However, animals can get to the same outcome in alternative ways. Thus, the outcome of the experiments does not tell us how animals achieve their performance.  In order to overcome this limitation, we used a set of computational models, which provide predictions on how alternative cognitive mechanisms perform in the face of varying environmental conditions, and compare their predictions to performance data from individuals that experienced different conditions throughout their life. Our study system is the cleaner fish *Labroides dimidiatus*, a coral reef fish that feeds off ectoparasites and dead tissue from the skin of other reef fishes. Cleaner fish often have to choose among alternative clients seeking their cleaning service, and the optimal choice is often hard to achieve. Our performance experiments mimic the natural choices the cleaner fish make. The combination of experiments and computational models points to cleaner fish abilities to account for the future effect of their choices by estimating the long term expected value of those choices. Estimating long term value is a mechanism involved in human foresight.

bibliography: Cleanerlearning.bib
header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{booktabs}
  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
output: rticles::plos_article
# plos.csl can be download and used locally
csl: http://www.zotero.org/styles/plos
---

_Text based on plos sample manuscript, see <https://journals.plos.org/ploscompbiol/s/latex>_

@@

Often alternative cognitive mechanisms yield similar behaviour and/or cognitive
performances. This poses a problem for disentangling the mechanistic 
underpinnings of behaviour. This is particularly clear in research aimed at
discovering between species variation in *higher* cognitive abilities; or 
in other words, research on whether non-human animals show cognitive 
abilities believed to be uniquely human. For instance, when researchers 
try to find *mental time travel-like* behaviour, they usually come up with 
clever experiments to show the behaviour displayed requires inferences made 
through past events[@dally_Foodcaching_2006]. However, they often face the
challenge of alternative scenarios where simpler explanations, 
like classic associative learning, can bring about the 
observed behavioural outcome [@suddendorf_Evolution_2007]. Similarly, attempts 
to demonstrate the presence of *theory of mind* in non-human animals 
face objections justified by alternative mechanisms underpinning 
similar behavioural results [@heyes_Theory_1998]. 
Such controversies are usually settled by using the principle 
of parsimony and its cognitive version, 
Lloyd Morgan’s cannon, which states that the simpler explanation (mechanism)
should be accepted. Ideally, however, alternative hypotheses should be evaluated 
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask between and within 
species variation that arises among different environments. Thus, if 
predictions from the alternative mechanisms depend on environmental 
conditions, it is possible to assess their relative explanatory power 
by performing cognitive test in the context of
different environmental conditions.

From an evolutionary perspective, mechanisms are likely selected because 
of how they allow individuals to respond to environmental variation. 
For example, biological market theory predicts that the exchange rate of goods 
and/or services traded between cooperative partners adjusts to the 
law of supply and demand, when individuals have some degree of partner choice
[@noe_Biological_1995]. Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should favour the ability 
to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviour to the varying conditions. More specifically the question arises 
to what extent mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station in their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering the same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish has eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally referred as the ephemeral reward task. 
Individuals are assessed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. 
This between variation in performance, and primates' known cognitive abilities
[@reader_Evolution_2001], shows that performance in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

Recent research on the cognitive abilities needed to solve the ephemeral 
reward task has used a broad approach as proposed by Shettleworth 
[@shettleworth_Cognition_2009], who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates  [@wismer_Cuebased_2019] while others find 
it in the food [@pretot_Comparative_2021; @pretot_Comparing_2016]. 
However, identifying cues as salient is not the only challenge 
of the ephemeral reward task, as revealed by proximate learning models. 
Such models allow varying the cognitive tool kit and evaluating 
which minimal kit is necessary to solve 
the task at hand (e.g. [@dubois_Model_2021]). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish face not only resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value 
one stimulus differently alone from compound versions of it has been termed 
configurational learning, chunking, 
or segmentation (see references in [@prat_Modelling_2022]).

In addition to configurational learning, cleaners also need to account for 
the future consequences of current decisions. In the model by 
Quiñones *et al.* [@quinones_Reinforcement_2019], this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behaviour as a negative reinforcer is, in principle, 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, the standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, however, the two 
mechanisms should be evaluated in light of how well 
they explain the available data. Note that different fields interested in
cognition and decision making use different words to refer to negative 
reinforcers [@quinones_Reinforcement_2019;@sutton_Reinforcement_2018]. 
Here, for the sake of simplicity and clarity, we will use the word 
'penalty' to refer to this mechanism which includes a negative reinforcer.

Over the last decade, our team tested over a hundred wild-caught 
cleaner fish in the exact same paradigm of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaner fish should not give priority to visitors any more.

@@

The model consists of a set of individual-based simulations where
individuals, representing cleaner fish, face a series of choices 
between two options, which simulate the natural conditions 
of the cleaning market. Individuals experience a series 
of discrete time points in which they face different ‘states’, 
defined by the number and category of client fish (visitor or residents) 
inviting for cleaning services. There are six possible states: 
zero clients, one resident, one visitor, 
resident-resident, visitor-visitor, and resident-visitor. 
The probability of each state is largely determined by the relative 
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination. 
This is because residents are willing to queue for cleaning service; 
while visitors leave the queue (with a certain probability) when made to wait. 
Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expected reward and the obtained
reward (prediction error) [@sutton_Reinforcement_2018; 
@rescorla_Theory_1972]. Given the new information, the update is carried
in the direction that leads to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor 
over a resident converges in the model. To which probability the model 
will converge depends on the relative abundance of cleaners, visitors 
and residents; as well as on the probability of visitors leaving the 
cleaning station when made to wait. Further details of the model 
implementation can be found in Quiñones *et al* [@quinones_Reinforcement_2019].

The model shows that agents need to find a way 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. 
If $\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases, individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards of an action (chaining).
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitors as a negative component (penalty). Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

@@

The aim of the analyses is to fit the key model parameters 
$\gamma$ and $\eta$, to the empirical data from Triki *et al.* 
[@triki_Biological_2019;@triki_Brain_2020] to test whether each or 
a combination of these effects is a better explanation for the pattern 
seen in the data. We used the ecological variables: cleaners, visitor clients, 
resident clients abundances and visitor clients leaving probability,  
as input to the models. As the response variable, we used cleaners' 
preference for visitors over residents in the ephemeral reward task. 
Finally, we used the preference for visitors resulting from the 
model simulations as the prediction for the response variable. 
We kept all other parameter values used for the model simulations constant, 
see Table \ref{tab:param}.

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors, we needed to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because, in the model, relative abundances of clients 
define not only the probability of residents and visitors 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable, which in nature may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of the site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances, we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

@@

To compare the fit of the three alternative models, we used the 
distribution of pseudo $R^2$ proposed by Mcfadden [@mcfadden_Conditional_1974].
Mcfadden's pseudo $R^2$ is a standard measure of fit for logistic regression. 
In that context, $pseudo-R^2$ uses the log-likelihood of the data given 
the model, relative to the log-likelihood of the data given a 
model without covariates, as a measure of fit. Our 
model is not a logistic regression, therefore we measured the $pseudo-R^2$ in
relation to the log-likelihood of a model
with parameters $\gamma$ and $\eta$ set to zero. This, in practice,
amounts to a model that has a neutral preference between the two options. 
We computed the pseudo $R^2$ for all the samples of the posterior from
the MCMC. Thus, we used these distributions of pseudo $R^2$'s as a measure of
fit.

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, penalty, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the rule these fish applied 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitors clients and generalized this rule to the lab conditions 
once being familiar with the task.

>>>>

---
title: | 
  Title: Field and experimental data together with computational models reveal how cleaner fish adjust decisions  in a biological market
# author:
# - Andrés E. Quiñones^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland - andresqp@gmail.com]
# - Zegni Triki^[Department of Zoology, Stockholm University, Stockholm, Sweden]
# - Redouan Bshary^[Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland]
subtitle: |
  Abbreviated title: Cleaner fish cognitive mechanisms in a biological market
# date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  word_document
   # pdf_document:
    # latex_engine: xelatex
    # template: template.latex
fig_caption: yes
fontsize: 12pt
keep_tex: yes
bibliography: Cleanerlearning.bib
style: M:\Zotero\styles\behavioral-ecology.csl
spacing: double
abstract: |
  While it is generally straightforward to quantify individual performance in cognitive experiments, identifying the underlying cognitive processes remains a major challenge. Often, different mechanistic underpinnings yield similar performances, and Lloyd Morgan’s cannon warrants acceptance of the simpler explanation. Alternatively, when the different mechanisms interact with environmental conditions, variation in performance across environments might allow to statistically infer the mechanism responsible. We illustrate this point by fitting computational models to experimental data on performance by wild-caught cleaner fish *Labroides dimidiatus* in an ephemeral reward task, as well as cleaner and client fish densities from the locations of capture. Using Bayesian statistics to fit the model parameters to performance data revealed that cleaner fish most likely estimate future consequences of an action, while it appears unlikely that the removal of the ephemeral reward acts as psychological punishment (negative reinforcement). Incorporating future consequences also yields performances that can be considered the result of locally optimal decision-rules, in contrast to the negative reinforcement mechanism. We argue that the combination of computational models with data is a powerful tool to infer the mechanistic underpinnings of cognitive performance.

header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{graphicx}
  \usepackage{booktabs}
  \linespread{2}
  \usepackage{lineno}
  \linenumbers
  \usepackage{setspace}\doublespacing
  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  \providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}
  ```
editor_options:
  markdown:
    wrap: 80
---

## Lay summary

  Cleaner fish, reef fish that feed of ectoparasites of other fishes, often choose among alternative clients. For the cleaner fish is better to choose certain clients, a choice that favours obtaining food in the future. Thus, cleaners must "look" into the future to make the right choice. By combining a learning model with data, we show that cleaners estimate the long-term value of their actions rather than using simpler heuristics. Estimating long-term value is a mechanism involved in human foresight.

**keywords:** learning, behavior, cleaners, bayesian statisitics, behavioral mechanisms

@@

Often alternative cognitive mechanisms yield similar behavior and/or cognitive
performances. This poses a problem for disentangling the mechanistic 
underpinnings of behavior. This is particularly clear in research aimed at
discovering between species variation in *higher* cognitive abilities; or 
in other words, research on whether non-human animals show cognitive 
abilities believed to be uniquely human. For instance, when researchers 
try to find *mental time travel-like* behavior, they usually come up with 
clever experiments to show the behavior displayed requires inferences made 
through past events [@dally_Foodcaching_2006]. However, they often face the
challenge of alternative scenarios where simpler explanations, 
like classic associative learning, can bring about the 
observed behavioural outcome [@suddendorf_Evolution_2007]. Similarly, attempts 
to demonstrate the presence of *theory of mind* in non-human animals 
face objections justified by alternative mechanisms underpinning 
similar behavioural results [@heyes_Theory_1998]. 
Such controversies are usually settled by using the principle 
of parsimony and its cognitive version, 
Lloyd Morgan’s cannon, which states that the simpler explanation (mechanism)
should be accepted. Ideally, however, alternative hypotheses should be evaluated 
in light of their explanatory power. One potential dimension that could allow
the variation necessary among distinct cognitive mechanisms is environmental
variation. Controlled laboratory conditions often mask between and within 
species variation that arises among different environments. Thus, if 
predictions from the alternative mechanisms depend on environmental 
conditions, it is possible to assess their relative explanatory power 
by performing cognitive tests in the context of
different environmental conditions.

From an evolutionary perspective, mechanisms are likely selected because 
of how they allow individuals to respond to environmental variation. 
For example, biological market theory predicts that the exchange rate of goods 
and/or services traded between cooperative partners adjusts to the 
law of supply and demand, when individuals have some degree of partner choice
[@noe_Biological_1995a]. Supply and demand conditions, which typically 
depend on the abundance of the species involved, certainly vary in time 
and space. Therefore, natural selection should favour the ability 
to flexibly adjust decisions and behavioural output to 
current market conditions. Indeed, such adjustments have been repeatedly 
documented [@axen_Signalling_1996]. In animals, an obvious general candidate 
mechanism for the strategic adjustment is the cognitive machinery. However, 
it is not clear which cognitive mechanisms allow individuals to adjust their 
behaviors to the varying conditions. More specifically the question arises 
to what extent mechanisms beyond basic associative learning may be involved.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Client fish seek cleaner fish services at their territory 
(so-called “cleaning station”) and offer themselves as food patches 
to get their ectoparasites removed, which provides cleaners 
with food and clients with improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016;@demaire_Reduced_2020]. 
Given the capacity of some client fish to swim larger distances and 
access multiple cleaning stations while others access the only cleaning 
station in their territory, it is crucial to categorize clients as 
either “visitors” or “residents”, respectively. During cleaning interactions, 
a cleaner fish often face a choice between a visitor and a resident client 
seeking its cleaning services simultaneously. Visitors have the option 
to switch to another cleaner fish if being made to wait, while residents 
must wait for inspection. Indeed,
visitors have been observed to use their partner choice option in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a].

When researchers aimed at testing wild-caught cleaner fish, as 
well as individuals from other species, abilities 
to prioritize visitor over resident clients in a lab-based paradigm, 
they used Plexiglas plates of different colours and shapes offering the same 
amount of food as surrogates for visitor and resident clients.
One plate is made to behave like a resident, i.e. it remains until 
the cleaner fish has eaten off it. The other plate was made to behave 
like a visitor, i.e. it is removed if not inspected first. 
This laboratory paradigm originally inspired by the cleaning market 
of cleaner fish is more generally referred as the ephemeral reward task. 
Individuals are assessed in their capacity to develop a preference for the 
ephemeral option, which gives them access to two reward units.
Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Furthermore, adult cleaner fish outperformed various primates as well as rats and
pigeons in this original version of the ephemeral reward task
[@salwiczek_Adult_2012; @zentall_Early_2017]; yet the African grey
parrots solve this task as well [@pepperberg_Can_2014]. 
This between variation in performance, and primates' known cognitive abilities
[@reader_Evolution_2011], shows that performance in the task
cannot be readily explained by classifying species as having high and 
low cognitive abilities. Instead, this between species variation suggests
that solving the ephemeral reward task requires a set of cognitive 
mechanisms that are not present in all the tested taxa, or were not used 
under the experimental set-up.

Recent research on the cognitive abilities needed to solve the ephemeral 
reward task has used a broad approach as proposed by
@shettleworth_Cognition_2009, who defined cognition as including 
all ways in which animals acquire information 
through the senses, process, retain and decide to act on it. 
Indeed, perception of relevant cues is of major importance 
for individual performance. For example, some animals find relevant 
information in the plates  [@wismer_Cuebased_2019] while others find 
it in the food [@pretot_Comparative_2021; @pretot_Comparing_2016]. 
However, identifying cues as salient is not the only challenge 
of the ephemeral reward task, as revealed by proximate learning models. 
Such models allow varying the cognitive tool kit and evaluating 
which minimal kit is necessary to solve 
the task at hand (e.g. @dubois_Model_2021). Applied to the ephemeral 
reward task, learning models showed that basic reinforcement learning does 
not suffice to solve the ephemeral reward task [@prat_Modelling_2022;
@quinones_Reinforcement_2019]. 
This is particularly so when models assume the more complex natural 
situation in which cleaner fish face not only resident-visitor pairs but also 
visitor-visitor and resident-resident pairs, as well a resident alone or 
a visitor alone. To be able to give visitors priority over residents, 
cleaners need to be able to assess a client’s value separately for the 
three possible scenarios (alone, paired with a fish with the 
same strategic option, paired with a fish with the alternative strategic option) 
[@quinones_Reinforcement_2019]. The ability to distinguish and value 
one stimulus differently alone from compound versions of it has been termed 
configurational learning, chunking, 
or segmentation (see references in @prat_Modelling_2022).

In addition to configurational learning, cleaners also need to account for 
the future consequences of current decisions. In the model by 
@quinones_Reinforcement_2019, this could be achieved in 
two non-mutually exclusive ways: through low temporal discounting of 
future effects, also termed ‘chaining’ [@enquist_Power_2016]; and/or 
through perceiving a visitor client leaving as psychological punishment 
(i.e. as a negative reinforcer). Low temporal discounting is when 
individuals include in their valuation of an action the reward effects 
that this will have in the future. This is done by combining in a single 
valuation the reward obtained in the current time with all the reward that 
comes after, discounting for how far in the future reward 
is accrued. 'Chaining' the reward of these different time steps allows 
individuals to take actions that increase the long-term reward at the sacrifice
of short term considerations [@enquist_Power_2016]. Even though, 'chaining' 
can be readily implemented computationally in learning models 
[@enquist_Power_2016;@sutton_Reinforcement_2018], cognitively it seems to be 
a complex adaptation [@suddendorf_Evolution_2007]. On the other hand,
using client behavior as a negative reinforcer is, in principle, 
easier to implement. Negative reinforcement is part of the ubiquitous 
learning mechanism termed operant conditioning [@thorndike_Animal_1898; 
@skinner_Behavior_1938]. Thus, the standard logic of 
Lloyd Morgan’s cannon demands that operant conditioning as the simpler 
explanation is to be accepted by default. Ideally, however, the two 
mechanisms should be evaluated in light of how well 
they explain the available data. Note that different fields interested in
cognition and decision making use different words to refer to negative 
reinforcers [@quinones_Reinforcement_2019;@sutton_Reinforcement_2018]. 
Here, for the sake of simplicity and clarity, we will use the word 
'penalty' to refer to this mechanism which includes a negative reinforcer.

Over the last decade, over a hundred wild-caught cleaner fish have been tested
in the exact same paradigm of the ephemeral reward 
task [@salwiczek_Adult_2012;@wismer_Variation_2014;@triki_Decrease_2018;
@triki_Biological_2019;@triki_Brain_2020]. These fish often come from 
different reef locations around the study field site at Lizard Island, 
Great Barrier Reef in Australia. Further investigation of the local 
eco-sociological conditions revealed that cleaner and client 
fish population densities have a substantial impact on cleaner fish 
performance in the task. Cleaner fish from reef sites with relatively 
low densities were more likely to fail at solving the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The explanation for this pattern
is twofold. First, low cleaner fish population densities imply 
low supply and high demand for cleaning services. 
Under such conditions, there are fewer occasions under 
which failing to give priority to a visitor translates to an empty 
cleaning station. Second and also related to the state of the cleaning market,
visitor clients at low cleaner fish population densities 
are less likely to exert their partner choice options 
and are hence more likely to wait for inspection if not given cleaning 
priority [@triki_Brain_2020]. In the absence of partner choice 
cleaner fish should not give priority to visitors any more.

@@

The model consists of a set of individual-based simulations where
individuals, representing cleaner fish, face a series of choices 
between two options, which simulate the natural conditions 
of the cleaning market. Individuals experience a series 
of discrete time points in which they face different ‘states’, 
defined by the number and category of client fish (visitor or residents) 
inviting for cleaning services. There are six possible states: 
zero clients, one resident, one visitor, 
resident-resident, visitor-visitor, and resident-visitor. 
The probability of each state is largely determined by the relative 
abundance of cleaner fish, residents and visitors, but to some degree by
cleaner fish choices when it faces the resident-visitor combination. 
This is because residents are willing to queue for cleaning service; 
while visitors leave the queue (with a certain probability) when made to wait. 
Individuals obtain a fixed reward from cleaning a client fish
regardless of the category. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expected reward and the obtained
reward - the prediction error- [@sutton_Reinforcement_2018; 
@rescorla_Theory_1972]. Given the new information, the update is carried
in the direction that leads to more reward being obtained, given
the new information. In the long run, the probability of choosing a visitor 
over a resident converges in the model. To which probability the model 
will converge depends on the relative abundance of cleaners, visitors 
and residents; as well as on the probability of visitors leaving the 
cleaning station when made to wait. Further details of the model 
implementation can be found in @quinones_Reinforcement_2019.

The model shows that agents need to find a way 
to incorporate future consequences of current choices. In the model, 
this could be achieved with either of two parameters that could 
also work together.  First, $\gamma$ measures how
much individuals include future rewards in their decision updates. 
If $\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases, individuals include more the
reward obtained from the subsequent choices. That amounts to estimating
and using for decision making the future expected rewards of an action (chaining).
Second, $\eta$ measures how much individuals include in their reward the
fleeing behavior of visitors as a negative component (penalty). Both of
these parameters allow individuals to use in their estimates the future
effects of their choices.

@@

The aim of the analyses is to fit the key model parameters 
$\gamma$ and $\eta$, to the empirical data from Triki et al. 
[-@triki_Biological_2019;@triki_Brain_2020] to test whether each or 
a combination of these effects is a better explanation for the pattern 
seen in the data. We used the ecological variables: cleaners, visitor clients, 
resident clients abundances and visitor clients leaving probability,  
as input to the models. As the response variable, we used cleaners' 
preference for visitors over residents in the ephemeral reward task. 
Finally, we used the preference for visitors resulting from the 
model simulations as the prediction for the response variable. 
We kept all other parameter values used for the model simulations constant, 
see Table \ref{tab:param}.

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors, we needed to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because, in the model, relative abundances of clients 
define not only the probability of residents and visitors 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable influencing the station occupancy, which in nature 
may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of the site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances, we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

@@

To compare the fit of the three alternative models, we used the 
distribution of pseudo $R^2$ proposed by Mcfadden [@mcfadden_Conditional_1974].
Mcfadden's pseudo $R^2$ is a standard measure of fit for logistic regression. 
In that context, $pseudo-R^2$ uses the log-likelihood of the data given 
the model, relative to the log-likelihood of the data given a 
model without covariates, as a measure of fit. Our 
model is not a logistic regression, therefore we measured the $pseudo-R^2$ in
relation to the log-likelihood of a model
with parameters $\gamma$ and $\eta$ set to zero. This, in practice,
amounts to a model that has a neutral preference between the two options. 
We computed the pseudo $R^2$ for all the samples of the posterior from
the MCMC. We used these distributions of pseudo $R^2$'s as a measure of
fit.

@@

In this study, our main aim was to unravel which of two potential 
cognitive mechanisms, chaining of events, penalty, 
or their combination, best explains wild-caught cleaner fish 
performance in the ephemeral reward task, while accounting for their
ecological conditions. To evaluate the merits of each of these two mechanisms
separately and combined, we considered cleaner fish performance 
in the lab test to have its origin from the rule these fish applied 
in their natural environment. That is, individuals that 
solved the task already had a preference for 
visitor clients and generalized this rule to the lab conditions 
once being familiar with the task.
